---
title: "Data Visualisation with ggplot2 - Project"
output: html_notebook
---

# AfterWork Data Science: Data Visualisation with Python Project

## 1. Defining the Question

### a) Specifying the Data Analysis Question
Specify the reasearch question that you'll be answering. i.e. Provide strategy recommendations that will lead to revenue growth.

### b) Defining the Metric for Success
The solutions to the following questions will help us answer our research question:

* When is the best time of year to book a hotel room?
* When is the optimal length of stay in order to get the best daily rate?
* How will you know if a hotel was likely to receive a disproportionately high number of special requests?

### c) Understanding the context
Provide some background information....

### d) Recording the Experimental Design
Describe the steps/approach that you will use to answer the given question.

### e) Data Relevance
How relevant was the provided data?


## 2. Reading the Data

```{r}
# Load the data below
# --- 
install.packages("tidyverse")
install.packages("ggplot2")
library("tidyverse")
library("ggplot2")
library(dplyr)
```


```{r}
# --- 
# YOUR CODE GOES BELOW
hotel_booking_df <- read.csv("hotel_bookings.csv")
```

```{r}
# Checking the first 5 rows of data
# ---
# YOUR CODE GOES BELOW
 head(hotel_booking_df)
```

```{r}
# Checking the last 5 rows of data
# ---
# YOUR CODE GOES BELOW
# 
tail(hotel_booking_df)
```

```{r}
# Sample 10 rows of data
# ---
# YOUR CODE GOES BELOW
# 
sample_n(hotel_booking_df,10)
```

```{r}
# Checking number of rows and columns
# ---
# YOUR CODE GOES BELOW
#  
dim(hotel_booking_df)

```

```{r}
# Checking datatypes
# ---
# YOUR CODE GOES BELOW
# 
str(hotel_booking_df)
```

```{r}
glimpse(hotel_booking_df)
```

Record your general observations below:

Observation 
We have  118201 rows  and 32 columns  in the date set


Observation 2


is_canceled  column only has  0  and 1   as intergers

## 3. External Data Source Validation
The data is originally from the article Hotel Booking Demand Datasets, by Nuno Antonio, Ana Almeida, and Luis Nunes for Data in Brief, Volume 22, February 2019.


## 4. Data Preparation
### Performing Data Cleaning
```{r}
# Checking datatypes and missing entries of all the variables
# ---
# YOUR CODE GOES BELOW
# 
hotel_df <- hotel_booking_df %>% replace(.=="NULL", NA)
colSums(is.na(hotel_df))

```

We observe the following from our dataset:

Observation 1

column  company has the highest null values   

Observation 2
column agent has second highest null values 


```{r}
# Checking how many duplicate rows are there in the data
# ---
# YOUR CODE GOES BELOW
# 
hotel_df[!duplicated(hotel_df), ]
#tail(hotel_booking_df)
```

We choose to keep the duplicates because we don't have a unique identifier to actually proof that we have duplicates.

```{r}
# Checking if any of the columns are all null
# ---
# YOUR CODE GOES BELOW
# 
colSums(is.na(hotel_df))
```

We observe the following from our dataset:

Observation 1
Highest column with missing values is company



```{r}
# Checking if any of the rows are all null
# ---
# YOUR CODE GOES BELOW
# 
is.na(hotel_df)
```

We observe the following from our dataset:

Observation 1
Observation 2

```{r}
# Checking the correlation of the features through the use of 
# visualizations the correlation using heatmap
# ---
# YOUR CODE GOES BELOW
# 
```

We observe the following from our dataset:

Observation 1
Observation 2

```{r}
# Dropping company column because it has alot of missing values 
# and we won't need to answer any of our questions
# ---
# YOUR CODE GOES BELOW
# 
hotel_df = select(hotel_df, -c("company"))
head(hotel_df)

```

From the data variable description we see that the Distribution Channel categoricy that tells us about Booking distribution.

The term “TA” means “Travel Agents”
The term “TO” means “Tour Operators”
This allows us to fill the missing values in the agents column with TO

```{r}
# We replace the mising values i.e. for TO
# ---
# YOUR GOES BELOW
#
hotel_df$agent[is.na(hotel_df$agent)] <- "TO"

tail(hotel_df)
```

```{r}
# We drop rows where there is no adult, baby and child as 
# these records won't help us.
# ---
# YOUR GOES BELOW
#

hotel1_df = filter(hotel_df, adults != 0 )
hotel2_df = filter(hotel1_df, children != 0 )
hotel3_df = filter(hotel2_df,babies !=0)
hotel3_df
```

```{r}
# We replace missing children values with rounded mean value
# ---
# Hint i.e. use round()
# ---
# YOUR GOES BELOW

sum(is.na(hotel3_df))
```

```{r}
# Checking for missing values in the dataframe
# ---
# YOUR GOES BELOW
#
sum(is.na(hotel3_df))
```

```{r}
# Converting the datatypes of the following columns from float to integer
# i.e. children, company, agent
# ---
# YOUR GOES BELOW
# 
str(hotel3_df)

## no float datatype
```


## 5. Solution Implementation

### 5.a) Questions

```{r}
# 1. How many bookings were cancelled?
# ---
# Visualisation: Barplot

# Create the summary table
hotel3_df %>% 
  group_by(hotel) %>%
  summarize( is_canceled = sum( is_canceled))

# Plotting the chart
ggplot(hotel3_df, aes(x = hotel, y = is_canceled)) +
  geom_col(
    stat="identity", width=0.5, fill = "#0099f9"
  ) +
  labs(
    title = "Hotel cancellation ", 
    x = "Hotel", 
    y = "Cancellations"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size=16, family="Times New Roman"), 
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```


```{r}
# 2. What was the booking ratio between resort hotel and city hotel?
# ---
# Barplot of booking ratio between resort hotel and city hotel

ggplot(data = hotel3_df,
       aes(
         x = hotel,
         y = prop.table(stat(count)),
         fill = factor(is_canceled),
         label = scales::percent(prop.table(stat(count)))
       )) +
  geom_bar(position = position_dodge()) +
  geom_text(
    stat = "count",
    position = position_dodge(.9),
    vjust = -0.5,
    size = 3
  ) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Reservation Status by Hotel Type", x = "Hotel Type", y = "Count") +
  theme_classic() +
  scale_fill_discrete(
    name = "Reservation Status",
    breaks = c("0", "1"),
    labels = c("Not Cancelled", "Cancelled")
  )
```


```{r}
# 3. What was the percentage of booking for each year?
# ---
# 
ggplot(data = hotel3_df,
       aes(
         x = hotel,
         y = prop.table(stat(count)),
         fill = factor(is_canceled),
         label = scales::percent(prop.table(stat(count)))
       )) +
  geom_bar(position = position_dodge()) +
  geom_text(
    stat = "count",
    position = position_dodge(.9),
    vjust = -0.5,
    size = 3
  ) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Reservation Status by Hotel Type", x = "Hotel Type", y = "Count") +
  theme_classic() +
  scale_fill_discrete(
    name = "Reservation Status",
    breaks = c("0", "1"),
    labels = c("Not Cancelled", "Cancelled")
  )

```

```{r}
# 4. Which were the most busiest months for hotels?
# ---
# 
datatest1 <-
  hotel_df %>% 
  mutate(arrival_date_month = factor(
    arrival_date_month,
    levels = c(
      "January",
      "February",
      "March",
      "April",
      "May",
      "June",
      "July",
      "August",
      "September",
      "October",
      "November",
      "December"
    ),
    ordered = TRUE
  ))

ggplot(hotel3_df, aes(arrival_date_month, fill = factor(is_canceled))) +
  geom_bar() +
  geom_text(stat = "count",
            aes(label = ..count..),
            position = position_stack(vjust = 0.5)) +
  scale_fill_discrete(
    name = "Reservation Status",
    breaks = c("0", "1"),
    label = c("Not Cancelled", "Cancelled")
  ) +
  labs(title = "Reservation Status by Month", x = "Month", y = "Count") +
  scale_x_discrete(labels = month.abb)
```


```{r}
# 5. From which top 3 countries did most guests come from?
# --- 
# YOUR GOES BELOW
#
hotel_booking_df%>%
  ggplot(aes(x=country,fill=hotel))+
  geom_bar()

```

```{r}
# 6.a) How long do most people stay in hotels?
#   b) By city and resort? Separate the data by hotel
# ---
# 
```


```{r}
# 7. Which was the most booked accommodation type (Single, Couple, Family)?
# ---
#
ggplot(hotel_df, aes(adults, fill=hotel)) + geom_bar(stat='count', position=position_dodge())
```

### 5.b) Recommendations
From the above analysis, below are our recommendations:

People refer resort hotel over city hotel
August to november is low season

## 6. Challenging your Solution
In this step, we review our solution and implement approaches that could potentially provide a better outcome. In our case, we could propose the following question that wasn't answered in our solution because it couldn't have greatly contributed to our recommendation.

```{r}
# When should hotels provide special offers?
# ---
# YOUR GOES BELOW
# 
```

Our observations:
-
Special offer should be on october 

How does this observation tie to our solution?

recorded lowest intake 

## 7. Follow up questions
During this step, you rethink and propose other ways that you can improve your solution.

a). Did we have the right data? Yes
b). Do we need other data to answer our question? Yes (country did not have sufficient data)
c). Did we have the right question?  yes



